
---
title: "Prediction Project"
author: "Andres Toledo"
date: "June 19, 2015"
output: html_document
---
Load the caret package
---
1.) read csv files and save the in separate variable setting NA values to missing fields
```{r}
library(caret)
```
Training : 
```{r}
training = read.csv("pml-training.csv", header=T, na.strings=c("NA", "#DIV/0!"))
dim(training)
```
Testing : 
```{r}
testing = read.csv("pml-testing.csv", header=T, na.string=c("NA", "#DIV/0!"))
dim(testing)
```

2.) Remove fields with NA values to avoid outliers

```{r}
naTrain<-training[, apply(training, 2, function(x) !any(is.na(x)))] 
cleanTrain<-naTrain[,-c(1:8)]
cleanTest<-testing[,names(cleanTrain[,-52])]
dim(cleanTrain)
dim(cleanTest)
```

3.)partitions data into a 60% training set and a 40%testing set

New Training : 

```{r}
trainingData<-createDataPartition(y=cleanTrain$classe, p=0.60,list=F)
newTraining<-cleanTrain[trainingData,] 
dim(newTraining)
```
New Testing :

```{r}
newTest<-cleanTrain[-trainingData,]
dim(newTest)
```

4.)Cross Validation and Random Forests

Set the seed

```{r}
set.seed(10119)
```

Implement crossvalidation
```{r}
fit<-trainControl(method="cv", number=2, allowParallel=T, verbose=T)
```

Generates random forest using crossvalidation

```{r}
randforest<-train(classe~.,data=newTraining, method="rf", trControl=fit, verbose=F)
```

5.)Prediction

Predicts values on the cleaned test set

```{r}
predrandforest<-predict(randforest, newdata=newTest)
```

Show the confusion matrix to identify the accuracy,sensitivity and specificity of the predictor.
The matrix does not show great variation altering the crossvalidation parameters which is why a low number was chosen to increase computation speed.

```{r}
confusionMatrix(predrandforest, newTest$classe)
```

Predicts values on original test data

```{r}
pred<-predict(randforest, newdata=cleanTest)
pred
```